Relationships in the extremes and their influence on competition and coexistence

### 2021-09-24 paper/Paper.Rnw
word and reference

%***DAN: Jasmin to edit the below as follows:
%1) Topic sentence: Another potentially important reason to quatify the importance of ATA to species coexistence is that there may
%be reason to believe that climate change can alter ATAs of environmental variables implicated in coexistence, as follows. 
%2) Just adapt below text, maybe try to simplify a bit.
%3) This para could just focus on the mechanics of how ATAs change for two variables measured at different times. 
%***DAN: Keeping this sentence so Jasmin can put it somwhere, not here.
%Given that storage effect depends on the relationship between E and C, we question the consequences of EC relationships characterized by ATA, the presence of which may be increasing as climate change brings about more extreme climatic events. 
%***DAN: I am also having a hard time identifying the topic sentence of this paragraph, can you tell me which one it is, please,
%in a comment right under this one?
% Extreme events can have important ecological impacts for storage effect and coexistence by introducing ATAs. 
An additional driving reason to quantify the significance of ATAs for species coexistence is the potential for climate change to alter ATAs of environmental variables implicated in coexistence, as follows. 
Among the various implications of climate change, the most potentially influential to ATAs in the environment is the increase in intensity, duration, extent, and frequency of extreme climatic events \citep{Jentsch_2007, Meehl_2004, Ummenhofer_2017}, i.e., heat-waves, floods, hurricanes, fires, etc. 
Temporally extensive extreme events can have important ecological impacts for
competing populations where species growth and the maintenance of biodiversity is dependent on phenologically separated environmental signals. A temporally extensive event can equalize the phenologically separated environmental signals, making them more correlated. For example, consider two precipitations signals, one in early spring and one in late summer, that are usually different. If a long flood occurs, it's likely that both signals will be close to equal in that year. If extreme floods occur in more years than not, then the two signals will become more correlated in higher precipitation values, a right-tailed ATA (figure \ref{fig:noise}C). Similarly, if extreme droughts occur in more years than not, then the signals will be more correlated in the lower values, a left-tailed ATA (figure \ref{fig:noise}A). The same can be said about spatially separated signals and more spatially extensive extreme events. ATA of environmental signals can then introduce or change ATAs in the fitness or traits of species that depend on the signals and potentially affect their coexistence.
%***DAN:
%1) Topic sentence: 
To make the above ideas more concrete, we review a relevant example from the literature that may be
influenced as previously described. 
We introduce a simplified example of an Arizona desert annual plant community from \cite{Chesson_2012}. 
%2) Make it concrete - name the system, name the organisms, name the seasons. 
The community follows the criteria required for temporal storage effects to operate. Different species dominate and have high germination rates depending on rainfall timing. For example, Baileya multiradiata dominates if sufficient rainfall is reached in the early germination season. Eriogonum abertianum dominates if sufficient rainfall is reached in the late germination season. Competition is influenced by resource consumption of seed-producing adult plants. Dormant seeds are unaffected by competition and can persist in the community in an unfavorable environment. When rain fall dynamics temporally fluctuate from year to year, relative abundance and dominance of the desert annual community also fluctuate year to year. The result is a relationship between the desert annuals' environmental responses (germination) that is symmetrically associated...

>>>>

%***DAN: Jasmin to edit the below as follows:
%1) Topic sentence: Another potentially important reason to quatify the importance of ATA to species coexistence is that there may
%be reason to believe that climate change can alter ATAs of environmental variables implicated in coexistence, as follows. 
%2) Just adapt below text, maybe try to simplify a bit.
%3) This para could just focus on the mechanics of how ATAs change for two variables measured at different times. 
%***DAN: Keeping this sentence so Jasmin can put it somwhere, not here.
%Given that storage effect depends on the relationship between E and C, we question the consequences of EC relationships characterized by ATA, the presence of which may be increasing as climate change brings about more extreme climatic events. 
%***DAN: I am also having a hard time identifying the topic sentence of this paragraph, can you tell me which one it is, please,
%in a comment right under this one?
% Extreme events can have important ecological impacts for storage effect and coexistence by introducing ATAs. 
An additional driving reason to quantify the significance of ATAs for species coexistence is the potential for climate change to alter ATAs of environmental variables required in coexistence, as follows. 
Among the various implications of climate change, the most potentially influential to ATAs in the environment is the increase in intensity, duration, extent, and frequency of extreme climatic events \citep{Jentsch_2007, Meehl_2004, Ummenhofer_2017}, i.e., heat-waves, floods, hurricanes, fires, etc. 
Temporally extensive extreme events can have important ecological impacts for
competing populations where species growth and the maintenance of biodiversity is dependent on phenologically separated environmental signals. A temporally extensive event can equalize the phenologically separated environmental signals, making them more correlated. For example, consider two precipitations signals, one in early spring and one in late summer, that are usually different. If a long flood occurs, it's likely that both signals will be close to equal in that year. If extreme floods occur in more years than not, then the two signals will become more correlated in higher precipitation values, a right-tailed ATA (figure \ref{fig:noise}C). Similarly, if extreme droughts occur in more years than not, then the signals will be more correlated in the lower values, a left-tailed ATA (figure \ref{fig:noise}A). The same can be said about spatially separated signals and more spatially extensive extreme events. ATA of environmental signals can then introduce or change ATAs in the fitness or traits of species that depend on the signals and potentially affect their coexistence.
%***DAN:
%1) Topic sentence: 
To make the above ideas more concrete, we review a relevant example from the literature that may be
influenced as previously described. 
We introduce a simplified example of an Arizona desert annual plant community from \cite{Chesson_2012}. 
%2) Make it concrete - name the system, name the organisms, name the seasons. 
The community follows the criteria required for temporal storage effects to operate. Different species dominate and have high germination rates depending on rainfall timing. For example, Baileya multiradiata dominates if sufficient rainfall is reached in the early germination season. Eriogonum abertianum dominates if sufficient rainfall is reached in the late germination season. Competition is influenced by resource consumption of seed-producing adult plants. Dormant seeds are unaffected by competition and can persist in the community in an unfavorable environment. When rain fall dynamics temporally fluctuate from year to year, relative abundance and dominance of the desert annual community also fluctuate year to year. The result is a relationship between the desert annuals' environmental responses (germination) that is symmetrically associated...
